---
title: "MATH 216 Data Science Final Project (Fall 2016)"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

---
output: 
  html_document: 
    keep_md: yes 
---


* Name: Katherine Hobbs
* Project title: Smoking Prevalence, Regulation, and Outcomes Across the United States
* Write-up on RPubs: <http://rpubs.com/khobbs/WriteUp>
* Abstract: 

The anti-tobacco campaign in the mid to late 20th century is widely considered one of the most successful public health campaigns in American history. Even so, according to the CDC, smoking remains the number one cause of preventable death and disease in the United States, causing roughly 480,000 deaths per year. This exploration of 2013 smoking and demographic data investigates the relationships between smoking prevelance, regulation, and outcomes in the United States. Results show that lung cancer death rates correlate positively with the percentage of smokers in a state and with cigarette consumption, while they correlate negatively with median household income. Additionally, while there is a very weak negative correlation between tobacco tax and lung cancer death rate, there is a much stronger negative correlation between tobacco tax and percentage of smokers, suggesting, but certaintly not proving, that tobacco taxes contribute to lower lung cancer death rates by detering people from smoking. 

```{r, echo=FALSE, message=FALSE}

library(tidyverse)
library(USAboundaries)
library(maptools)
library(sp)
library(broom)
library(leaflet)
library(dplyr)
library(ggplot2)
library(rgdal)
library(readr)
library(forcats)
library(knitr)
library(dtplyr)

#Load Data Sets

Lung_Cancer <- read_csv("Lung Cancer Social Explorer.csv")

Taxes <- read_csv("Tax Data.csv")

Tobacco_Use <- read_csv("Tobacco Use.csv")

State_Income <- read_csv("Income.csv")

#Clean Data Sets

Tobacco_Use <- Tobacco_Use %>% 
  select(`Location Description`,`Data Value`, Gender) %>% 
  spread("Gender", "Data Value")

Taxes <- Taxes %>% 
  select(`Location Description`,`Sub-Measure`,`Data Value`) %>% 
  spread("Sub-Measure", "Data Value", fill= NA)

#Rename variables

Lung_Cancer <- Lung_Cancer %>% 
  rename(LC_Death_Rate = SE_T038_001) %>% 
  rename(LC_Deaths = SE_NV121_001) %>% 
  rename(State=Geo_NAME)

Lung_Cancer <- Lung_Cancer %>% 
  select(State, LC_Deaths, LC_Death_Rate)

State_Income <- State_Income %>% 
  rename(population=SE_T001_001) %>% 
  rename(median_hs_income=SE_T057_001) %>% 
  rename(State=Geo_NAME)

State_Income <- State_Income %>% 
  select(State, population, median_hs_income)

Taxes <- Taxes %>% 
  rename(State=`Location Description`) %>% 
  rename(Avg_Pack_Cost=`Average Cost per pack`) %>% 
  rename(Consumption=`Cigarette Consumption (Pack Sales Per Capita)`) %>% 
  rename(Total_Tax_percent=`Federal and State tax as a Percentage of Retail Price`) %>% 
  rename(Total_Tax_amount=`Federal and State Tax per pack`) %>% 
  rename(State_Tax_amount=`State Tax per pack`)

Taxes <- Taxes %>% 
  select(State, Avg_Pack_Cost, Consumption, Total_Tax_percent, Total_Tax_amount)
  
Tobacco_Use <- Tobacco_Use %>% 
  rename(State=`Location Description`) %>% 
  rename(Current_Smokers_All=`Overall`) %>% 
  rename(Current_Smokers_Male=`Male`) %>% 
  rename(Current_Smokers_Female=`Female`)

#Filter out territories and non-state entries

Tobacco_Use <- Tobacco_Use %>% 
  filter(State!="Puerto Rico") %>% 
  filter(State!="Guam") %>% 
  filter(State!="National Median (States and DC)") %>% 
  filter(State!="District of Columbia")

Taxes <- Taxes %>% 
  filter(State!="District of Columbia")

Lung_Cancer <- Lung_Cancer %>% 
  filter(State!="Puerto Rico") %>% 
  filter(State!="District of Columbia")

#Join the Lung_Cancer, State_Income, Taxes, and Tobacco_Use datasets by State

join_1 <- left_join(Taxes, Tobacco_Use, by="State")
join_2 <- left_join(join_1, Lung_Cancer, by="State")
Complete_State_Data <- left_join(join_2, State_Income, by="State")

#Add number of smokers variable

Complete_State_Data <- Complete_State_Data %>% 
  mutate(Numb_Smokers=population*Current_Smokers_All)

#Load US States Polygons Data
us_states <- us_states() 
us_states_tidy <- tidy(us_states, region="geoid")

#Load US States Metadata
us_states_data <- us_states@data

#Join US States Metadata
us_states_combined <- left_join(us_states_tidy, us_states_data, by=c("id"="geoid"))
us_states_combined <- us_states_combined %>% 
  filter(name!="Puerto Rico") %>% 
  filter(name!="District of Columbia")

#Join Combined States Data with Complete_State_Data
Map_Data <- left_join(us_states_combined, Complete_State_Data, by=c("name"="State"))

ggplot(data=Map_Data, aes(x=long, y=lat, group=group, fill=LC_Death_Rate)) +
  geom_polygon() +
  scale_fill_gradient(low="white", high="darkblue") + 
  geom_path(size=0.1) +
  coord_map(xlim=c(-130, -65),ylim=c(23,50)) +
  ggtitle("Lung Cancer Death Rate By State")
```

